The effects of context inference on motor learning: savings, de-adaptation and spontaneous recovery

### 2021-07-19 adaptation/article/article.tex
Update on Overleaf.

\subsection{Context inference}
In our model, the idea of context is quite general. The context comprises the relevant elements of hte environment (e.g. high winds when walking), as well as the task at hand. Previous models have focused on the elements of the context which are directly related to the internal models: given a past action, the different internal models (e.g. $f_B$ and $f_H$ from the previous section) make different predictions for the outcome of that action; the model that best predicts that action is deemed to be the ``correct'' model, i.e. the model that best represents the context.

@@

In the previous section, we used the variable $\vartheta$ to refer to the outcomes of actions. In this section, we expand the definition of an observation $\vartheta$ to include any information given to the decision-making agent by the environment. This includes, as before, the outcomes of actions, but also any other sensory information that might be indirectly related to actions, or even completely intependent from them. For this section, the most important component of $\vartheta$ is contextual information, i.e. information that might directly help infer the identity of the current context.

@@

Note that, because context inference integrates information from many sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use. For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone. Proprioceptive signals can also provide contextual information [EGs of experiments with hidden hands]. The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than normal trial-by-trial variation in performance.

@@

Our model offers a different explanation for the observed results: the assymetry is due to the existance of the baseline context. As discussed above, action selection in our model is done via a weighted avarage:
\begin{equation}
p(a_t | s_t) \propto \displaystyle \sum_{m \in M} p(m | s_t, s_{t-1}, a_{t-1}) p(a_t | m)
\end{equation}
In experiments without cued contexts, the baseline model $m_O$ has a non-zero probability $p(m | s_t ...)$. When a new block of trials starts (e.g. in the transition from 3A to A), a switch is inferred by the model (given feedback after the first trial) and $m_O$ becomes more likely (given that $m_{3A}$ has been ruled out. Therefore, in these first trials, action selection has a component guided by the baseline model, in which no extra compensatory force is applied, essentially ``pulling'' adaptation towards zero (no compensatory force). In the first group, this initial pull towards zero accelerates the transition towards A, as $3A > A > 0$, but on the second group, it slows down the switch because $A > 0 > -A$.

>>>>

\subsection{Context inference}
In our model, the idea of context is quite general. The context comprises the relevant elements of the environment (e.g. high winds when walking), as well as the task at hand. Previous models have focused on the elements of the context which are directly related to the internal models: given a past action, the different internal models (e.g. $f_B$ and $f_H$ from the previous section) make different predictions for the outcome of that action; the model that best predicts that action is deemed to be the ``correct'' model, i.e. the model that best represents the context.

@@

In the previous section, we used the variable $\vartheta$ to refer to the outcomes of actions. In this section, we expand the definition of an observation $\vartheta$ to include any information given to the decision-making agent by the environment. This includes, as before, the outcomes of actions, but also any other sensory information that might be indirectly related to actions, or even completely independent from them. For this section, the most important component of $\vartheta$ is contextual information, i.e. information that might directly help infer the identity of the current context.

@@

Note that because context inference integrates information from many sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use. For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone. Proprioceptive signals can also provide contextual information [EGs of experiments with hidden hands]. The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than normal trial-by-trial variation in performance.

@@

Our model offers a different explanation for the observed results: the asymmetry is due to the existence of the baseline context. As discussed above, action selection in our model is done via a weighted average:
\begin{equation}
p(a_t | s_t) \propto \displaystyle \sum_{m \in M} p(m | s_t, s_{t-1}, a_{t-1}) p(a_t | m)
\end{equation}
In experiments without cued contexts, the baseline model $m_O$ has a non-zero probability $p(m | s_t ...)$. When a new block of trials starts (e.g. in the transition from 3A to A), a switch is inferred by the model (given feedback after the first trial) and $m_O$ becomes more likely (given that $m_{3A}$ has been ruled out. Therefore, in these first trials, action selection has a component guided by the baseline model, in which no extra compensatory force is applied, essentially ``pulling'' adaptation towards zero (no compensatory force). In the first group, this initial pull towards zero accelerates the transition towards A, as $3A > A > 0$, but on the second group, it slows down the switch because $A > 0 > -A$.
